Reanalysis Step 1: Data Preparation

### 2025-08-14 reanalysis-1-data_preparation.qmd
update

@@

    # Comorbidity indicators as factors
    glaucoma_bef_hitplus_cat = factor(if_else(Glaucoma_bef_hitplus == "Present", 1, 0)),
    otherretina_bef_hitplus_cat = factor(if_else(Otherretina_bef_hitplus == "Present", 1, 0)),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus),

@@

```{r missing-assessment}
# Variables used in propensity score model
ps_model_vars <- c(
  "ever_lapse_binary",
  "baseline_VA_logMAR",
  "gender_cat",
  "race_ethnic_cat", 
  "insurance_cat",
  "age_cat",
  "CCI",
  "DCSI",
  "other_inject",
  "anti_VEGF",
  "focal_laser_flag",
  "PRP_flag",
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat",
  "catsurg_before_hitplus_cat"
)

@@

# Assess missing data for analysis variables
missing_summary <- df_analysis |>
  dplyr::select(dplyr::all_of(c(ps_model_vars, "outcome_va_vi_binary"))) |>
  dplyr::summarise(dplyr::across(everything(), ~ sum(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to = "variable", values_to = "n_missing") |>
  dplyr::mutate(
    pct_missing = round(n_missing / nrow(df_analysis) * 100, 2),
    var_type = dplyr::case_when(
      variable %in% c("outcome_va_vi_binary") ~ "Outcome",
      variable == "ever_lapse_binary" ~ "Exposure",
      TRUE ~ "Covariate"
    )
  ) |>
  dplyr::arrange(desc(n_missing))

# Display missing data summary
missing_summary |> 
  filter(n_missing > 0) |>
  kable(caption = "Variables with missing data (for imputation)")
```

>>>>

## Missing Data Handling Strategy

**Following the original analysis approach:**

1. **Glaucoma/Other Retina Conditions**: Missing values are recoded as 0 (absence of condition)
   - Clinical assumption: These conditions would be documented if present
   - This recoding happens BEFORE propensity score modeling

2. **CCI/DCSI**: Rows with missing values will be excluded (complete case)
   - These are key confounders that cannot be reliably imputed

3. **Other variables**: Missing values will be handled via multiple imputation in the next step

@@

    # Comorbidity indicators as factors (missing = no condition)
    glaucoma_bef_hitplus_cat = factor(case_when(
      Glaucoma_bef_hitplus == "Present" ~ 1,
      TRUE ~ 0  # Includes NA values coded as 0 (no glaucoma)
    )),
    otherretina_bef_hitplus_cat = factor(case_when(
      Otherretina_bef_hitplus == "Present" ~ 1,
      TRUE ~ 0  # Includes NA values coded as 0 (no other retina condition)
    )),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus),

@@

```{r missing-assessment}
# Variables used in propensity score model
ps_model_vars <- c(
  "ever_lapse_binary",
  "baseline_VA_logMAR",
  "gender_cat",
  "race_ethnic_cat", 
  "insurance_cat",
  "age_cat",
  "CCI",
  "DCSI",
  "other_inject",
  "anti_VEGF",
  "focal_laser_flag",
  "PRP_flag",
  "glaucoma_bef_hitplus_cat",  # Missing already coded as 0 (no condition)
  "otherretina_bef_hitplus_cat",  # Missing already coded as 0 (no condition)
  "catsurg_before_hitplus_cat"
)

@@

# Assess missing data BEFORE recoding (for documentation)
missing_before_recoding <- df_combined |>
  dplyr::select(dplyr::all_of(c(
    "baseline_VA_logMAR", "CCI", "DCSI", 
    "Glaucoma_bef_hitplus", "Otherretina_bef_hitplus",
    "cataract_surgery_bef_hitplus"
  ))) |>
  dplyr::summarise(dplyr::across(everything(), ~ sum(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to = "variable", values_to = "n_missing_original")

# Assess missing data AFTER recoding
missing_summary <- df_analysis |>
  dplyr::select(dplyr::all_of(c(ps_model_vars, "outcome_va_vi_binary"))) |>
  dplyr::summarise(dplyr::across(everything(), ~ sum(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to = "variable", values_to = "n_missing") |>
  dplyr::mutate(
    pct_missing = round(n_missing / nrow(df_analysis) * 100, 2),
    var_type = dplyr::case_when(
      variable %in% c("outcome_va_vi_binary") ~ "Outcome",
      variable == "ever_lapse_binary" ~ "Exposure",
      variable %in% c("glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat") ~ "Covariate (recoded)",
      TRUE ~ "Covariate"
    ),
    handling = dplyr::case_when(
      variable %in% c("CCI", "DCSI") ~ "Exclude if missing",
      variable %in% c("glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat") ~ "Missing → 0 (no condition)",
      TRUE ~ "To be imputed"
    )
  ) |>
  dplyr::arrange(desc(n_missing))

# Display missing data summary
missing_summary |> 
  filter(n_missing > 0) |>
  kable(caption = "Variables with missing data after recoding (glaucoma/retina conditions coded as 0)")
```
